---
title: "Workflow: Data Scientist Creating Accessible Dashboards"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow: Data Scientist Creating Accessible Dashboards}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
set.seed(42)
```

## The Scenario

You're a data scientist creating a dashboard to visualize patient outcomes across 
treatment groups. Your requirements are:

- **6 distinct colors** for different treatment groups
- **CVD-safe** for accessibility compliance  
- **Include company brand colors** (a specific blue and orange)
- **Easy integration** with ggplot2

This vignette shows how huerd makes this workflow simple and reliable.

## Quick Start: The Simple Path

For many use cases, `quick_palette()` is all you need:
```{r}
library(huerd)

# Generate a 6-color accessible palette
palette <- quick_palette(6)
print(palette)
```

Want to include your brand colors? Just add them:

```{r}
brand_palette <- quick_palette(
  n = 6,
  brand_colors = c("#1f77b4", "#ff7f0e")  # Your blue and orange
)
print(brand_palette)
```

## Using Palettes with ggplot2

huerd integrates natively with ggplot2, so you no longer need `scale_color_manual()`:

```{r}
library(ggplot2)

# Simulate treatment outcome data
set.seed(123)
treatment_data <- data.frame(
  patient_id = 1:180,
  treatment = rep(c("Control", "Drug A", "Drug B", "Drug C", "Combo 1", "Combo 2"), each = 30),
  outcome = c(
    rnorm(30, 50, 10),
    rnorm(30, 55, 10),
    rnorm(30, 60, 10),
    rnorm(30, 58, 10),
    rnorm(30, 65, 10),
    rnorm(30, 62, 10)
  ),
  time_weeks = rep(1:30, 6)
)

# Plot with automatic huerd palette
ggplot(treatment_data, aes(x = time_weeks, y = outcome, color = treatment)) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_color_huerd() +
  labs(
    title = "Patient Outcomes by Treatment Group",
    x = "Weeks",
    y = "Outcome Score",
    color = "Treatment"
  ) +
  theme_minimal()
```

### With Brand Colors

Include your brand colors directly in the scale:

```{r}
ggplot(treatment_data, aes(x = time_weeks, y = outcome, color = treatment)) +
  geom_smooth(method = "loess", se = FALSE) +
  scale_color_huerd(brand_colors = c("#1f77b4", "#ff7f0e")) +
  labs(
    title = "Patient Outcomes by Treatment Group",
    subtitle = "Using brand colors",
    x = "Weeks",
    y = "Outcome Score",
    color = "Treatment"
  ) +
  theme_minimal()
```

### Using a Pre-Generated Palette

For consistent colors across multiple plots, generate once and reuse:

```{r}
# Generate and save your palette
my_palette <- generate_palette(
  n = 6,
  include_colors = c("#1f77b4", "#ff7f0e"),
  progress = FALSE
)

# Use in multiple plots
p1 <- ggplot(treatment_data, aes(x = treatment, y = outcome, fill = treatment)) +
  geom_boxplot() +
  scale_fill_huerd(palette = my_palette) +
  labs(title = "Distribution by Treatment") +
  theme_minimal() +
  theme(legend.position = "none")

p2 <- ggplot(treatment_data, aes(x = outcome, fill = treatment)) +
  geom_density(alpha = 0.6) +
  scale_fill_huerd(palette = my_palette) +
  labs(title = "Outcome Density") +
  theme_minimal()

# Display side by side
gridExtra::grid.arrange(p1, p2, ncol = 2)
```

## Verifying Accessibility

Before finalizing your dashboard, verify accessibility:

### Quick Check

```{r}
is_cvd_safe(my_palette)
```

### Detailed Assessment

```{r}
interpret_palette_quality(my_palette)
```

### Full Evaluation

```{r}
evaluation <- evaluate_palette(my_palette)

# Key metrics for your accessibility report
cat("Minimum color distance:", round(evaluation$distances$min, 3), "\n")
cat("CVD worst-case distance:", round(evaluation$cvd_safety$worst_case_min_distance, 3), "\n")
cat("Performance ratio:", round(evaluation$distances$performance_ratio * 100, 1), "%\n")
```

## Visual Diagnostics

For presentations or documentation, use the analysis dashboard:

```{r, fig.width=10, fig.height=6}
plot_palette_analysis(my_palette)
```

Or the simple swatch view:

```{r, fig.width=7, fig.height=2}
plot(my_palette)
```

## Summary

For data scientists, huerd provides:

1. **`quick_palette()`** - Fast, sensible defaults
2. **`scale_color_huerd()` / `scale_fill_huerd()`** - Native ggplot2 integration
3. **`is_cvd_safe()`** - Quick accessibility verification
4. **`interpret_palette_quality()`** - Human-readable quality assessment
5. **`plot_palette_analysis()`** - Full visual diagnostics

This workflow ensures your visualizations are beautiful and accessible.
