---
title: "Workflow: Designer Creating Brand-Compliant Palettes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow: Designer Creating Brand-Compliant Palettes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
set.seed(2024)
```

## The Scenario

You're a designer creating a color palette for a client's annual report. Your requirements are:

- **Include specific brand colors** that the client requires
- **Generate complementary colors** that work with the brand
- **Ensure accessibility** for compliance requirements
- **Export to multiple formats** (CSS, Sass, etc.)
- **Document and justify** your color choices

This vignette shows how huerd simplifies brand-compliant palette creation.

## Starting with Brand Colors

Use `brand_palette()` when you have specific colors that must be included:

```{r}
library(huerd)

# Client's brand colors
client_navy <- "#003366"
client_orange <- "#FF6600"

# Create a 6-color palette around the brand
palette <- brand_palette(
  brand_colors = c(client_navy, client_orange),
  n_total = 6
)

print(palette)
```

The brand colors are preserved exactly, and huerd generates optimized complementary colors.

## Visualizing Your Palette

See your palette at a glance:

```{r, fig.width=7, fig.height=2}
plot(palette)
```

For client presentations, use the full analysis dashboard:

```{r, fig.width=10, fig.height=6}
plot_palette_analysis(palette)
```

## Controlling Lightness for Print

For print materials, you may want to avoid very light or very dark colors:

```{r}
# Mid-range lightness for better print reproduction
print_palette <- quick_palette(
  n = 6,
  brand_colors = c(client_navy, client_orange),
  lightness = "mid"  # Avoids extremes
)

print(print_palette)
```

Or specify custom bounds:

```{r}
# Very specific lightness range
controlled_palette <- generate_palette(
  n = 6,
  include_colors = c(client_navy, client_orange),
  init_lightness_bounds = c(0.3, 0.7),  # 30% to 70% lightness
  progress = FALSE
)
```

## Documenting for Stakeholders

### Human-Readable Quality Assessment

```{r}
quality <- interpret_palette_quality(palette)
print(quality)
```

This gives you language you can use directly in client presentations.

### Technical Metrics

For detailed documentation:

```{r}
evaluation <- evaluate_palette(palette)

cat("=== Palette Quality Report ===\n\n")
cat("Number of colors:", evaluation$n_colors, "\n")
cat("Minimum perceptual distance:", round(evaluation$distances$min, 3), "\n")
cat("Mean perceptual distance:", round(evaluation$distances$mean, 3), "\n")
cat("Optimization performance:", round(evaluation$distances$performance_ratio * 100, 1), "%\n")
cat("\n=== Accessibility ===\n")
cat("CVD worst-case distance:", round(evaluation$cvd_safety$worst_case_min_distance, 3), "\n")
cat("CVD-safe:", is_cvd_safe(palette), "\n")
```

## CVD Simulation for Compliance

Show stakeholders how the palette appears to colorblind viewers:

```{r}
cvd_sim <- simulate_palette_cvd(palette, cvd_type = "all")
print(cvd_sim)
```

## Exporting Your Palette

huerd can export to multiple formats for development handoff:

### CSS Custom Properties

```{r}
css_output <- export_palette(
  palette,
  format = "css",
  names = c("brand-navy", "brand-orange", "accent-1", "accent-2", "accent-3", "accent-4")
)
cat(css_output)
```

### Sass Variables

```{r}
sass_output <- export_palette(
  palette,
  format = "sass",
  names = c("brand-navy", "brand-orange", "accent-1", "accent-2", "accent-3", "accent-4")
)
cat(sass_output)
```

### JSON for Web Applications

```{r}
json_output <- export_palette(palette, format = "json")
cat(json_output)
```

### Save to Files

```{r, eval=FALSE}
# Export directly to files
export_palette(palette, format = "css", file = "brand-colors.css")
export_palette(palette, format = "sass", file = "_brand-colors.scss")
export_palette(palette, format = "json", file = "brand-colors.json")
```

## Reproducibility for Version Control

Every palette includes metadata for exact reproduction:

```{r}
# Original palette
original <- generate_palette(
  n = 6,
  include_colors = c(client_navy, client_orange),
  progress = FALSE
)

# Reproduce exactly (e.g., in a different session)
reproduced <- reproduce_palette(original, progress = FALSE)

# Verify they're identical
identical(as.character(original), as.character(reproduced))
```

## Complete Design Workflow Example

```{r, fig.width=10, fig.height=6}
# 1. Define brand colors
brand_colors <- c("#003366", "#FF6600")

# 2. Generate palette
final_palette <- brand_palette(brand_colors, n_total = 6)

# 3. Verify quality
quality_report <- interpret_palette_quality(final_palette)
print(quality_report)

# 4. Check accessibility
cat("\nAccessibility Check:", if(is_cvd_safe(final_palette)) "PASS" else "REVIEW NEEDED", "\n")

# 5. Visual review
plot_palette_analysis(final_palette)

# 6. Export for development
cat("\n=== CSS Export ===\n")
cat(export_palette(final_palette, format = "css",
                   names = c("primary", "secondary", "accent1", "accent2", "accent3", "accent4")))
```

## Summary

For designers, huerd provides:

1. **`brand_palette()`** - Start with your required brand colors
2. **`quick_palette(lightness = ...)`** - Control lightness for print/digital
3. **`interpret_palette_quality()`** - Client-ready quality language
4. **`simulate_palette_cvd()`** - Accessibility documentation
5. **`export_palette()`** - CSS, Sass, JSON, CSV exports
6. **`reproduce_palette()`** - Exact reproducibility for version control

This workflow ensures your brand palettes are optimized, accessible, and properly documented.
