---
title: "Introduction to huerd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to huerd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(42)
```

## Introduction

The `huerd` package provides a scientifically-grounded approach to generating categorical color palettes. It uses a pure minimax optimization algorithm in the perceptually uniform OKLAB color space to create palettes with maximally distinct colors.

This vignette will walk you through the core features of `huerd`, from basic palette generation to in-depth analysis.

## Installation

You can install the development version of `huerd` from GitHub with:

```R
# install.packages("remotes")
# remotes::install_github("sims1253/huerd")
```

## Basic Palette Generation

The simplest way to use `huerd` is with the `generate_palette()` function. By default, it will create a palette of the specified size with colors that are as distinct as possible.

```{r}
library(huerd)

# Generate a palette of 5 colors
palette <- generate_palette(5, progress = FALSE)
print(palette)
```

All palettes are automatically sorted by brightness (lightness in the OKLAB space), making them intuitive to use.

## Constrained Palettes

A key feature of `huerd` is the ability to include fixed "brand" colors in your palette while optimizing the remaining colors around them.

```{r}
# Generate a 6-color palette that must include a specific blue and orange
brand_palette <- generate_palette(
  n = 6,
  include_colors = c("#4A6B8A", "#E5A04C"),
  progress = FALSE
)
print(brand_palette)
```

## Palette Analysis

`huerd` includes powerful tools for analyzing the quality of your palettes.

### `evaluate_palette()`

The `evaluate_palette()` function provides a detailed, quantitative assessment of a palette's properties.

```{r}
# Evaluate the brand palette we just created
evaluation <- evaluate_palette(brand_palette)
print(evaluation)
```

This function returns a wealth of information, including:

*   **Perceptual Distances**: Minimum, mean, and other statistics about the distances between colors.
*   **CVD Safety**: How the palette performs under simulated color vision deficiency.
*   **Color Distribution**: Statistics on the spread of lightness, chroma, and hue.

### `plot_palette_analysis()`

For a more visual analysis, the `plot_palette_analysis()` function creates a comprehensive dashboard.

```{r, fig.width=10, fig.height=6}
# Create the diagnostic dashboard
plot_palette_analysis(brand_palette)
```

This dashboard provides six key visualizations:

1.  **Color Swatches**: An overview of the palette with key metrics.
2.  **Pairwise Distance Matrix**: A heatmap showing the perceptual distance between every pair of colors.
3.  **Nearest Neighbor Distances**: A bar chart showing how distinct each color is from its closest neighbor.
4.  **OKLAB Color Space**: A projection of the colors in the `a*b*` plane of the OKLAB space.
5.  **CVD Simulation**: How the palette appears to individuals with the three most common types of color vision deficiency.
6.  **Comparative Palettes**: A comparison of your palette's distance distribution against established palettes like Viridis and Set2.

## CVD Accessibility

`huerd` provides two main tools for working with CVD.

### `is_cvd_safe()`

This function provides a simple, programmatic check to see if a palette meets a minimum threshold for CVD safety.

```{r}
is_cvd_safe(brand_palette)
```

### `simulate_palette_cvd()`

This function allows you to see how your palette would appear to individuals with different types of CVD.

```{r}
# Simulate the appearance for all CVD types
cvd_simulation <- simulate_palette_cvd(brand_palette, cvd_type = "all")
print(cvd_simulation)
```

## Conclusion

This vignette has covered the core functionality of the `huerd` package. By combining pure minimax optimization with comprehensive analysis tools, `huerd` provides a powerful and flexible solution for creating high-quality, accessible color palettes.
