---
title: "Introduction to huerd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to huerd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
set.seed(42)
```

## Introduction

The `huerd` package generates categorical color palettes with pure minimax optimization in the perceptually uniform OKLAB color space, so the colors are maximally distinct.

This vignette covers the core features of `huerd`, from basic palette generation to in-depth analysis.

## Installation

You can install the development version of `huerd` from GitHub with:

```R
# install.packages("remotes")
# remotes::install_github("sims1253/huerd")
```

## Basic Palette Generation

The simplest way to use `huerd` is with the `generate_palette()` function. By default, it creates a palette of the specified size with colors as distinct as possible.

```{r}
library(huerd)

# Generate a palette of 5 colors
palette <- generate_palette(5, progress = FALSE)
print(palette)
```

huerd automatically sorts every palette by brightness (OKLAB lightness).

## Constrained Palettes

You can include fixed "brand" colors in your palette while `huerd` optimizes the remaining colors around them.

```{r}
# Generate a 6-color palette that must include a specific blue and orange
brand_palette <- generate_palette(
  n = 6,
  include_colors = c("#4A6B8A", "#E5A04C"),
  progress = FALSE
)
print(brand_palette)
```

## Palette Analysis

`huerd` includes tools for analyzing the quality of your palettes.

### `evaluate_palette()`

`evaluate_palette()` gives a detailed, quantitative assessment of a palette's properties.

```{r}
# Evaluate the brand palette we just created
evaluation <- evaluate_palette(brand_palette)
print(evaluation)
```

This function returns detailed information, including:

*   **Perceptual Distances**: Minimum, mean, and other statistics about the distances between colors.
*   **CVD Safety**: How the palette performs under simulated color vision deficiency.
*   **Color Distribution**: Statistics on the spread of lightness, chroma, and hue.

### `plot_palette_analysis()`

For a more visual analysis, `plot_palette_analysis()` creates a diagnostic dashboard.

```{r, fig.width=10, fig.height=6}
# Create the diagnostic dashboard
plot_palette_analysis(brand_palette)
```

This dashboard provides six key visualizations:

1.  **Color Swatches**: An overview of the palette with key metrics.
2.  **OKLAB Color Space**: A projection of the colors in the `a*b*` plane; point size indicates lightness.
3.  **Pairwise Distance Matrix**: A heatmap of the perceptual distance between every pair of colors, on a fixed scale so cell colors are comparable across palettes.
4.  **CVD Simulation**: How the palette appears to viewers with the three most common types of color vision deficiency, plus greyscale.
5.  **Pairwise Distances under CVD**: The distribution of pairwise distances under each CVD simulation, compared with normal vision.
6.  **Comparative Palettes**: A comparison of your palette's distance distribution against established palettes like Batlow, Viridis, and Set2.

## CVD Accessibility

`huerd` provides two main tools for working with CVD.

### `is_cvd_safe()`

This function checks programmatically whether a palette meets a minimum threshold for CVD safety.

```{r}
is_cvd_safe(brand_palette)
```

### `simulate_palette_cvd()`

This function shows how your palette would appear to viewers with different types of CVD.

```{r}
# Simulate the appearance for all CVD types
cvd_simulation <- simulate_palette_cvd(brand_palette, cvd_type = "all")
print(cvd_simulation)
```

## Conclusion

This vignette has covered the core functionality of the `huerd` package. By combining pure minimax optimization with its analysis tools, `huerd` creates high-quality, accessible color palettes.
