---
title: "Workflow: Package Developer Integrating huerd"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow: Package Developer Integrating huerd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The Scenario

You're developing an R package that includes visualization functions. You want to:

- **Generate optimal palettes** automatically based on data
- **Ensure reproducibility** for scientific workflows
- **Handle edge cases** gracefully
- **Provide quality metrics** to users
- **Minimize dependencies** and runtime overhead

This vignette demonstrates how to integrate huerd into your package.

## Adding huerd as a Dependency

In your package's `DESCRIPTION`:

```
Imports:
    huerd
```
Or for optional functionality:

```
Suggests:
    huerd
```

## Basic Integration Pattern

### Automatic Palette Generation

```{r}
library(huerd)

# Function that automatically generates palettes
plot_clusters <- function(data, cluster_column, ...) {
  n_clusters <- length(unique(data[[cluster_column]]))
  
  # Generate optimal palette for this number of clusters
  palette <- generate_palette(
    n = n_clusters,
    progress = FALSE  # Silent for programmatic use
  )
  
  # Use the palette...
  list(
    n_clusters = n_clusters,
    palette = palette
  )
}

# Example usage
result <- plot_clusters(iris, "Species")
print(result$palette)
```

### Reproducible Palettes

For scientific workflows, ensure reproducibility:

```{r}
# Set seed before generation
set.seed(42)
palette1 <- generate_palette(8, progress = FALSE)

# Same seed = same palette
set.seed(42)
palette2 <- generate_palette(8, progress = FALSE)

identical(as.character(palette1), as.character(palette2))
```

Or use the built-in reproduction:

```{r}
original <- generate_palette(8, progress = FALSE)
reproduced <- reproduce_palette(original, progress = FALSE)

identical(as.character(original), as.character(reproduced))
```

## Accessing Optimization Metadata

huerd palettes include rich metadata for logging and debugging:

```{r}
palette <- generate_palette(
  n = 8,
  return_metrics = TRUE,  # Include quality metrics
  progress = FALSE
)

# Access optimization details
opt_details <- attr(palette, "optimization_details")
str(opt_details)

# Access quality metrics
metrics <- attr(palette, "metrics")
cat("Minimum distance:", metrics$distances$min, "\n")
cat("Performance ratio:", metrics$distances$performance_ratio, "\n")
```

## Choosing the Right Optimizer

Different optimizers suit different needs:

```{r}
# COBYLA (default) - Good balance, deterministic
cobyla_pal <- generate_palette(6, optimizer = "nloptr_cobyla", progress = FALSE)

# SANN - Stochastic, may find better solutions
sann_pal <- generate_palette(6, optimizer = "sann", progress = FALSE)

# L-BFGS - Fast gradient-based, best with smooth objectives
lbfgs_pal <- generate_palette(
  6,
  optimizer = "nlopt_lbfgs",
  weights = c(smooth_repulsion = 1),
  progress = FALSE
)

# Compare results
cat("COBYLA min distance:", evaluate_palette(cobyla_pal)$distances$min, "\n")
cat("SANN min distance:", evaluate_palette(sann_pal)$distances$min, "\n")
cat("L-BFGS min distance:", evaluate_palette(lbfgs_pal)$distances$min, "\n")
```

**Recommendations:**
- Use `nloptr_cobyla` (default) for general use
- Use `sann` when quality is more important than speed
- Use `nlopt_lbfgs` with `smooth_repulsion` for fastest optimization

## Handling Edge Cases

### Single Color

```{r}
single <- generate_palette(1, progress = FALSE)
print(single)
length(single)
```

### Empty Palette

```{r}
empty <- generate_palette(0, progress = FALSE)
print(empty)
length(empty)
```

### Large Palettes

```{r}
# For many colors, consider limiting iterations for speed
large <- generate_palette(
  n = 15,
  max_iterations = 500,  # Faster but potentially lower quality
  progress = FALSE
)
print(large)
```

## Evaluating External Palettes

Your package might want to compare huerd palettes against user-provided ones:

```{r}
# Evaluate any vector of hex colors
viridis_colors <- c("#440154", "#3B528B", "#21918C", "#5DC863", "#FDE725")

eval_viridis <- evaluate_palette(viridis_colors)
cat("Viridis (5 colors):\n")
cat("  Min distance:", round(eval_viridis$distances$min, 3), "\n")
cat("  CVD safety:", round(eval_viridis$cvd_safety$worst_case_min_distance, 3), "\n")

# Compare with huerd
huerd_colors <- generate_palette(5, progress = FALSE)
eval_huerd <- evaluate_palette(huerd_colors)
cat("\nhuerd (5 colors):\n")
cat("  Min distance:", round(eval_huerd$distances$min, 3), "\n")
cat("  CVD safety:", round(eval_huerd$cvd_safety$worst_case_min_distance, 3), "\n")
```

## Palette Objects as Character Vectors

huerd palettes work seamlessly as character vectors:

```{r}
palette <- generate_palette(5, progress = FALSE)

# Standard vector operations
length(palette)
palette[1:3]
rev(palette)
paste(palette, collapse = ", ")

# Coercion
as.character(palette)

# Works with base R graphics
barplot(1:5, col = palette)
```

## Integration with ggplot2

If your package uses ggplot2, huerd provides native scales:

```{r, eval=requireNamespace("ggplot2", quietly = TRUE)}
library(ggplot2)

# Automatic palette
ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  scale_color_huerd()

# Pre-generated palette for consistency
my_palette <- generate_palette(3, progress = FALSE)

ggplot(iris, aes(Sepal.Length, Sepal.Width, color = Species)) +
  geom_point() +
  scale_color_huerd(palette = my_palette)
```

## Conditional huerd Usage

If huerd is in `Suggests`, check for availability:

```{r, eval=FALSE}
my_plot_function <- function(data, ...) {
  n_groups <- length(unique(data$group))
  
  if (requireNamespace("huerd", quietly = TRUE)) {
    palette <- huerd::generate_palette(n_groups, progress = FALSE)
  } else {
    # Fallback to default colors
    palette <- grDevices::hcl.colors(n_groups, "Set2")
  }
  
  # Use palette...
}
```

## Performance Considerations

```{r}
# Benchmark different configurations
library(huerd)

# Fast generation (exploration)
# Expected runtime: ~0.1-0.3 seconds (low iterations for quick exploration)
system.time({
  fast <- generate_palette(8, max_iterations = 100, progress = FALSE)
})

# Default generation
# Expected runtime: ~0.5-2.0 seconds (balanced iteration count for good quality)
system.time({
  default <- generate_palette(8, progress = FALSE)
})

# High quality generation
# Expected runtime: ~2.0-8.0 seconds (high iterations for maximum quality)
system.time({
  quality <- generate_palette(8, max_iterations = 3000, progress = FALSE)
})
```

For performance-critical applications:

1. **Cache palettes** - Generate once, reuse across plots
2. **Use `max_iterations`** - Lower values for faster (but potentially lower quality) results
3. **Use `nlopt_lbfgs`** - Fastest optimizer for smooth objectives

## Summary

For package developers, huerd provides:

1. **Programmatic API** - `progress = FALSE` for silent operation
2. **Reproducibility** - Seed control and `reproduce_palette()`
3. **Rich metadata** - `attr(palette, "optimization_details")` for logging
4. **Multiple optimizers** - Choose speed vs. quality
5. **Edge case handling** - Works with n=0, n=1, large n
6. **Standard R integration** - Works as character vector
7. **ggplot2 scales** - Native `scale_color_huerd()` integration
8. **External evaluation** - `evaluate_palette()` works on any colors

This makes huerd suitable for both interactive use and integration into automated pipelines.
